Estimating duration in the presence of misclassification

### 2024-02-20 biology-data.tex
ch5 proof
fix links
add all intro stuff needed

\subsubsection{Strata}

@@

\subsubsection{Positivity results}

@@

\subsubsection{Episodes}

@@

\todo[inline]{%
    From chapter 7 add info on: how this is aggregated/stratified/etc, SRS computational limitations, definitions for each episode (sketched below).
}

The beginning of episode $i$ is known to occur in the interval $[l_i^{(b)}, r_i^{(b)}]$, and similarly for the end of the infection in $[l_i^{(e)}, r_i^{(e)}]$.

>>>>

\subsubsection{Strata} \label{biology-data:sec:cis-strata}

@@

\subsubsection{Positivity results} \label{biology-data:sec:positivity-results}

@@

\subsubsection{Episodes} \label{biology-data:sec:cis-episodes}

@@

The crucial features of an infection episode are the start and end times of the infection.
Ignoring misclassified test results, latent start time of infection episode $i$, $B_i$, can be bounded as between the day after the last negative test, $l_i^{(b)}$, and the day of the first positive test, $r_i^{(b)}$.
Similarly, the latent end time of infection episode $i$, $E_i$, can be bounded as between the day of the last positive test, $l_i^{(e)}$, and the day before the following negative test, $r_i^{(e)}$.

\subsubsection{Computational limitations and privacy protection}

For privacy reasons, the CIS data is stored in the ONS's SRS (Secure Research Service).
The SRS is a Trusted Research Environment (TRE) that allows researchers to access sensitive data in a secure environment.

Statistical disclosure controls must be applied to any data or results to be extracted from the SRS.
Statistical disclosure control ensures that no individual can be identified from the data.
This means that data must be aggregated to counts of at least 5, and no individual-data included.

Analyses requiring more fine-grained stratification (such as that in \cref{biology-data:sec:MRP}) or linking data from the same individual over time (such as that in \cref{E-perf-test,E-imperf-test}) must be performed within the SRS.
Therefore, computationally efficient approaches must be taken to this work.

\section{Conclusion}

In this chapter, I have introduced RT-PCR testing, the natural history of SARS-CoV-2, and the two studies used in this thesis.
The next chapter will introduce the data generating processes behind infection, prevalence, and our observations of these processes.
Then, I will apply these concept to the two studies.

### 2024-02-20 incidence-prevalence.tex
ch5 proof
fix links
add all intro stuff needed

\emph{Duration}, or \emph{time-to-event}, distributions can be expressed in several equivalent ways~\autocite[17]{sunStatisticala}.
The duration distribution I will be concerned with is the distribution of the number of days for which an individual is prevalent.
Consider an individual $i$ infected on day $B_i$ that is prevalent during the interval $[B_i, E_i]$.
\todo{check $B_i$ and $E_i$ are not defined elsewhere}
I define this period as an \emph{infection episode}, beginning on day $B_i$ and ending on day $E_i$.
$D_i = E_i - B_i + 1$ is their \emph{duration of positivity}, the number of days for which they are prevalent.
I model the $D_i$s as a discrete independent and identically distributed (iid) random variable; extending the methodology to account for covariates is left to further work.
The duration is a positive integer, \ie $D_i \in \{1, 2, \dots, \dmax \}$, where $\dmax$ is the maximum duration.
The first way of expressing the distribution of $D_i$ is as the standard probability mass function (pmf), $f_{D}(d) = \prob(D = d)$; or cumulative density function (CDF), $F_{D}(d) = \prob(D \leq d) = \sum_{i=1}^d f_{D}(i)$.
The \emph{survival function} gives the probability that the duration is at least $d$ days: $S(d) = \prob(D \geq d) = 1 - F_{D}(D - 1)$ with $S(1) = 1$.
%(a common alternative convention for survival analysis with discrete variables is to define the survival function as $\prob(D >d)$).
The \emph{hazard} on day $d$ is the probability of recovering on day $d$ conditional on still being positive on day $d$: $\lambda(d) = \prob(D = d \mid D \geq d) = f_{D}(d) / S(d)$.
Throughout this thesis, the survival function will be the primary quantity of interest, and it will normally be expressed in terms of the hazard; these two quantities are related by $S(d) = \prod_{i=1}^{d-1} (1 - \lambda(i))$.

@@

% Numerical methods for solving systems of this form, that is solving for $\vec{Z}$ (non-random in this expression) for a known $\E (\vec{P} \mid \vec{Z} )$, are well-established~\autocite[e.g.:][section 8.2]{highamAccuracy}.
% However, these algorithms can be unstable, meaning that numerical errors are large.
% This is a particular issue if $S(1)$ is small.
% Since, by definition, $S(1)=1$ here, this is not an issue in this context.
%However, if instead of estimating the incidence of RT-PCR positives, the aim was to estimate the incidence of infections, then $S(1)$ would be small and this would be an issue.
%While this quantity is more epidemiologically relevant, the time from infection to RT-PCR positivity is short and, for now, I will neglect the difference.

@@

Robust, unbiased prevalence estimates already exist (\cref{E-biology-data:sec:positivity}), and incidence estimates are the objective of this thesis.
This chapter has shown that the key quantity relating these quantities is the distribution of the duration of RT-PCR positivity.
However, no estimates of this quantity in the general population exist\todo{ref lit review of duration}.
Therefore, I now turn to estimating the duration of RT-PCR positivity.

>>>>

\emph{Duration}, or \emph{time-to-event}, distributions can be expressed in several equivalent ways~\autocite[17]{sunStatisticala}.
The duration distribution I will be concerned with is the distribution of the number of days for which an individual is prevalent.
Consider an individual $i$ infected on day $B_i$ that is prevalent during the interval $[B_i, E_i]$ (as in \cref{E-biology-data:sec:cis-episodes}).
I define this period as an \emph{infection episode}, beginning on day $B_i$ and ending on day $E_i$.
$D_i = E_i - B_i + 1$ is their \emph{duration of positivity}, the number of days for which they are prevalent.
I model the $D_i$s as a discrete independent and identically distributed (iid) random variable; extending the methodology to account for covariates is left to further work.
The duration is a positive integer, \ie $D_i \in \{1, 2, \dots, \dmax \}$, where $\dmax$ is the maximum duration.
The first way of expressing the distribution of $D_i$ is as the standard probability mass function (pmf), $f_{D}(d) = \prob(D = d)$; or cumulative density function (CDF), $F_{D}(d) = \prob(D \leq d) = \sum_{i=1}^d f_{D}(i)$.
The \emph{survival function} gives the probability that the duration is at least $d$ days: $S(d) = \prob(D \geq d) = 1 - F_{D}(D - 1)$ with $S(1) = 1$.
%(a common alternative convention for survival analysis with discrete variables is to define the survival function as $\prob(D >d)$).
The \emph{hazard} on day $d$ is the probability of recovering on day $d$ conditional on still being positive on day $d$: $\lambda(d) = \prob(D = d \mid D \geq d) = f_{D}(d) / S(d)$.
Throughout this thesis, the survival function will be the primary quantity of interest, and it will normally be expressed in terms of the hazard; these two quantities are related by $S(d) = \prod_{i=1}^{d-1} (1 - \lambda(i))$.

@@

% Numerical methods for solving systems of this form, that is solving for $\vec{Z}$ (non-random in this expression) for a known $\E (\vec{P} \mid \vec{Z} )$, are well-established~\autocite[e.g.][section 8.2]{highamAccuracy}.
% However, these algorithms can be unstable, meaning that numerical errors are large.
% This is a particular issue if $S(1)$ is small.
% Since, by definition, $S(1)=1$ here, this is not an issue in this context.
%However, if instead of estimating the incidence of RT-PCR positives, the aim was to estimate the incidence of infections, then $S(1)$ would be small and this would be an issue.
%While this quantity is more epidemiologically relevant, the time from infection to RT-PCR positivity is short and, for now, I will neglect the difference.

@@

Robust, unbiased prevalence estimates already exist (\cref{E-biology-data:sec:cis}), and incidence estimates are the objective of this thesis.
This chapter has shown that the key quantity relating these quantities is the distribution of the duration of RT-PCR positivity.
However, no estimates of this quantity in the general population exist\todo{ref lit review of duration}.
Therefore, I now turn to estimating the duration of RT-PCR positivity.
